Errors in Landcover Data Limit Our Understanding of Global Change

### 2016-03-13 paper/main.tex
Manuscript revised for GCB format

%% PNAStwoS.tex
%% Sample file to use for PNAS articles prepared in LaTeX
%% For two column PNAS articles
%% Version1: Apr 15, 2008
%% Version2: Oct 04, 2013

%% BASIC CLASS FILE
\documentclass{pnastwo2}

%% ADDITIONAL OPTIONAL STYLE FILES Font specification

%\usepackage{pnastwoF}
%\usepackage{cite}
\usepackage[numbers,round,sort,compress]{natbib}
\bibpunct{(}{)}{,}{n}{,}{,}  % https://xianblog.wordpress.com/tag/natbib/ (allows natbib with PNAS)
\usepackage{lineno}
\setlength\linenumbersep{2pt}
%\usepackage{placeins}
%\usepackage{float}
%\restylefloat*{figure}
%\usepackage{dblfloatfix}  % http://tex.stackexchange.com/questions/235623/placing-a-figure-in-the-bottom-of-a-page-spanning-the-two-columns-of-an-ieee-doc 
%\usepackage{subfigure}
%\usepackage{wrapfig}
%\usepackage{caption}
%\setlength\belowcaptionskip{-2cm}
%\setlength{\textfloatsep}{5pt plus 1.0pt minus 2.0pt}
\usepackage{multirow}
\setlength{\parskip}{1cm plus4mm minus3mm}
\usepackage{makecell}
\usepackage{boldline}
\setcellgapes{3pt}
\usepackage[flushleft]{threeparttable}

%% OPTIONAL MACRO DEFINITIONS
\def\s{\sigma}
%%%%%%%%%%%%
%% For PNAS Only:
%\url{www.pnas.org/cgi/doi/10.1073/pnas.0709640104}
\copyrightyear{2008}
\issuedate{Issue Date}
\volume{Volume}
\issuenumber{Issue Number}
%\setcounter{page}{2687} %Set page number here if desired
%%%%%%%%%%%%

\begin{document}
%\hspace*{\fill}

@@

\author{Lyndon Estes{\affil{1}{Woodrow Wilson School, Princeton University, Princeton, NJ USA}\affil{2}{Civil and Environmental Engineering, Princeton University, Princeton, NJ USA}},
Peng Chen\affil{3}{Indiana University, Bloomington, IN USA},
Stephanie Debats\affil{2}{},
Tom Evans\affil{3}{},
Stefanus Ferreira\affil{4}{GeoTerraImage, Pretoria, RSA}{},
Gabrielle Ragazzo\affil{5}{Ecology and Evolutionary Biology, Princeton University, Princeton, NJ USA},
Justin Sheffield\affil{2}{},
Adam Wolf\affil{5}{},
Eric Wood\affil{2}{},
\and
Kelly Caylor\affil{2}{}}

@@

\maketitle

\begin{article}
\begin{abstract}
{Research into global environmental change is often built upon landcover maps, which can have substantial errors. How these errors impact our understanding of global change processes and associated policies is thus of great interest, but is difficult to fully quantify because we lack spatially comprehensive reference data, particularly in the World's most rapidly developing regions. We used a high quality, national-scale cropland map to assess bias and accuracy in current generation landcover maps and four examples of ``downstream'' (landcover-dependent) analyses, two related to physical processes (estimates of vegetative carbon stocks and evapotranspiration) and two to socio-economic processes (gridded crop production estimates and agent-based simulation of household food security), at resolutions ranging from 1-100 km. We found that the landcover maps have substantial errors below 25 km resolution, with substantial biases and inaccuracies, particularly from maps derived from coarse resolution sensors. In some cases these error metrics remain high even when maps are aggregated to 100 km. These errors can be substantially amplified in downstream analyses (e.g. the carbon and crop production analyses), but in cases where the values of different cover types that underpin the analysis are similar (e.g. evapotranspiration estimates) the results can be relatively insensitive. To avoid these errors, and thus minimize the risk of misunderstanding global change processes or misinforming policy, substantial map aggregation of landcover maps from their base resolution is often needed. We provide several guidelines that can help landcover map users select appropriate datasets and  aggregation scales under different use cases.

@@

\keywords{landcover | bias | remote sensing | agriculture | crop yield | harvested area | carbon | agent-based model | landscape}

@@

\dropcap{T}he functioning of the Earth System is fundamentally connected to the characteristics of landcover \cite{lambin_modelling_1997}, the physical constituents of the terrestrial surface. Human endeavors are strongly governed by and shape landcover, whether it be felling forests for timber or burning savannas to rejuvenate forage, and landcover is a primary factor in many climate and biogeochemical processes \cite{lambin_dynamics_2003}. Our increasing modification of the Earth's surface \cite{lambin_dynamics_2003} means that socioeconomic and physical processes increasingly interact through landcover. To fully understand these processes and the nature of global change, it is essential to know the nature and distribution of landcover.

This importance is understood by a growing number of social, economic, and natural scientists, who are using landcover data to advance understanding of food security \cite{lark_cropland_2015,wright_recent_2013, licker_mind_2010}, carbon cycling \cite{asner_high-resolution_2010, gaveau_major_2014}, biodiversity loss \cite{newbold_global_2015, luoto_predicting_2004}, demographic shifts \cite{linard_assessing_2010}, and other important facets of global change.

The value of the insights resulting from such studies depends upon the veracity of their underlying landcover data, much as a house requires a solid foundation in order to remain standing. Unfortunately, the evidence to date indicates that much of our understanding of global change is built on shaky foundations. The reason for this is that landcover data can only practically be derived from satellite imaging, which has several important constraints--related to both physics and analytics--that propagate mapping errors. First, in many regions the cover types of interest are smaller than the sensor resolution \cite[e.g. smallholder's farms][]{jain_mapping_2013,debats_generalized_2015,ozdogan_resolution_2006}, or spectrally indistinct from neighboring covers \cite{sweeney_mapping_2015}, which are factors that increase mapping complexity \cite{yu_meta-discoveries_2014}. Second, the act of defining a cover class can cause error, in that selected classes may have highly diverse spectral properties (e.g. agriculture \cite{debats_generalized_2015,estes_platform_2015}) and thus be difficult for the classifier to distinguish, while discretizing a continuous cover type (e.g. dividing a forest into different canopy cover classes) can also promote classification error, particularly near class boundaries \cite{foody_status_2002}, as well as confusion about the actual extent of the cover type \cite{sexton_conservation_2015}. Furthermore, class definitions often vary between maps, complicating inter-comparison \cite{kuemmerle_challenges_2013}. Third, landcover map series are often used to detect changes \cite[e.g.][]{gross_monitoring_2013}, or may use within-class seasonality as a means for discriminating between classes (e.g. phenology of cropping versus surrounding vegetation \cite{sweeney_mapping_2015}), but seasonal variability and landcover changes can be easily confused. Given these multiple sources of error, landcover maps are often inaccurate at finer scales and disagree widely between products, particularly in the world's most rapidly developing regions \cite{estes_projected_2013, fritz_comparison_2010, fritz_cropland_2011}. These errors limit our ability to obtain granular, mechanistic understanding of global change processes.

These problems with landcover products is known \cite{fritz_comparison_2010, fritz_cropland_2011, see_improved_2015, fritz_mapping_2015,verburg_challenges_2011}, and there are a variety of map improvement efforts underway \cite[e.g.][]{fritz_geo-wiki:_2012, estes_platform_2015}. What remains an open question is exactly how much the maps researchers typically use deviate from actual landcover, and how this in turn impacts our understanding of global change processes. Fully answering this question depends on having comprehensive, spatially representative ground truth data, which are typically unavailable for Africa and other rapidly changing regions \cite{see_improved_2015}, which have also been fairly sparsely mapped \cite{yu_meta-discoveries_2014}. Our current understanding of map accuracy over such areas is therefore based on a small number of maps \cite{yu_meta-discoveries_2014}, which were assessed using either bottom-up tests made with a relatively small number of ground truth points, or from top-down "sanity checks" made in comparison to aggregated survey data.  In terms of the former, there are well-developed, best-practice guidelines for selecting and using ground-truth samples to robustly quantify map error, but these are not commonly applied by the broader community of landcover map users (often due to lack of data), which makes it challenging to compare how error varies within and between maps \cite{stehman_global_2012, olofsson_making_2013,olofsson_good_2014,foody_status_2002}. The latter form of assessment are made at coarser administrative levels relative to reported statistics \cite[e.g.][]{fritz_comparison_2010}, which can also be suspect \cite{carletto_emperor_2013,fao_action_2013}.

Since it is difficult to fully quantify landcover map errors, it is even more challenging to gauge their impact on downstream analyses, in which there is substantial risk of error amplification\cite{kuemmerle_challenges_2013}. There has been some work examining how map error influences climate simulations \cite{ge_impacts_2007}, agricultural land use patterns \cite{schmit_limitations_2006}, carbon flux measurements \cite{quaife_impact_2008}, human population estimates \cite{linard_assessing_2010}, and species distribution models \cite{tuanmu_global_2014}, but these either use simulated landcover errors \cite{ge_impacts_2007} or compare relevant differences in estimates between different satellite-derived landcover maps \cite{linard_assessing_2010, quaife_impact_2008,tuanmu_global_2014}. One exception is a Belgian study \cite{schmit_limitations_2006} that used ground-collected farm parcel data to assess how landcover errors bias measurements of agricultural land use patterns, but the study extent was fairly small and the validation data were discontiguous.

Fortunately, the recent, explosive growth in public and private initiatives to develop new Earth observing capabilities, which range from small drones\footnote{e.g. 3DRobotics, DJIA} to new high resolution satellite arrays\footnote{Planet Labs, Skybox} and better mapping methods \cite{fritz_geo-wiki:_2012,tuanmu_global_2014,estes_projected_2013, debats_generalized_2015}, are finally providing the means to more comprehensively interrogate the accuracy and biases in the landcover products that have become commonplace in global change research--and which are often used to make policy decisions \cite{searchinger_high_2015}.

In this study, we take advantage of this recent growth in high resolution data to address the call to more thoroughly assess errors and in landcover maps \cite{kuemmerle_challenges_2013, olofsson_good_2014,olofsson_global_2012}, and further examine how these errors might impact our understanding of change in the world's most dynamic regions.  We use a unique, high accuracy landcover map of South African crop fields to conduct a spatially comprehensive, bottom-up quantification of error in several widely used landcover maps, and how these errors can propagate through downstream studies investigating both the physical and socioeconomic components of global change. Our objective is to provide scientists and policy-makers who use landcover maps an up-to-date assessment of their appropriate uses and limitations.

\vspace{-0.5 cm}
\section{Overview of study area and analyses}
In the late 2000s, the South African government commissioned a cropland map that was made by manually interpreting and digitizing fields visible within high resolution satellite imagery \cite{fourie_better_2009}. The resulting vectorized field boundaries provide highly accurate data on field sizes and distribution for the period 2009-2011. Measured in terms of its ability to distinguish crop fields from other cover types, this dataset is 97\% percent accurate at 1 km resolution, with user's accuracies of 94\% and 98\% for the cropland and non-cropland classes, and producer's accuracies of 84 and 99\%  \cite[][SI]{estes_platform_2015}.

This dataset is particularly valuable because South Africa represents nearly 6\% of sub-Saharan Africa's (SSA) area, which is a region that is poised to undergo rapid agricultural expansion \cite{searchinger_high_2015}, yet is notably lacking trustworthy maps of existing agriculture land \cite{fritz_comparison_2010}. Moreover, South Africa's large agricultural sector represents the diversity of farming systems found throughout SSA, ranging from large commercial operations to smallholder farms \cite{hardy_rainfed_2011,estes_using_2014}.

We used this dataset as a reference layer for evaluating four landcover products representative of the type commonly used in global change studies. The first was South Africa's own 30 m resolution 2009 National Landcover map (SA-LC)\cite{sanbi_national_2009}, which is typical of the higher-resolution, Landsat-based maps that are typically available only for individual countries \cite[e.g.][]{fry_completion_2009}. The second and third were respectively the 300 m GlobCover 2009 \cite{arino_global_2012} and 500 m resolution 2011 MODIS Landcover products, which are widely used global-scale products \cite[e.g.][]{gross_monitoring_2013, shackelford_conservation_2015}. The fourth dataset was the new 1 km GeoWiki hybrid-fusion cropland map for Africa \cite{fritz_mapping_2015}, which incorporates the first three datasets and represents the current state-of-the-art in landcover mapping.  Since GeoWiki provides a continuous bounded value of landcover (percent cropland cover), which cannot be feasibly converted to a categorical value, we converted all other datasets, including the reference vectors, into comparable 1 km gridded percent cropland estimates.

We first evaluated the quality of the landcover product-derived percent cropland maps (hereafter referred to as the ``test maps''). Instead of the standard confusion matrix-based accuracy assessment metrics \cite{olofsson_good_2014,olofsson_making_2013}, which apply to categorical landcover maps, we assessed the bias and accuracy of the test maps based on the gridded residuals that resulted when each test map was subtracted from the reference map. Here bias is the mean residual value, weighted by the density of reference cropland, and accuracy is the mean of the absolute values of residuals (also weighted by cropland density), thus accuracy is inversely related to this value. We calculated these metrics for the original 1 km resolution, and for maps that were further aggregated to 5, 10, 25, 50, and 100 km, in order to evaluate how bias and accuracy changes with scale. We also assessed how landcover pattern impacts map performance by modeling the correlation between map accuracy and cropland density.

We then used these maps to conduct four further analyses typical of global change research. The first two related to physical processes, namely the calculation of vegetative carbon stocks and the simulation of evapotranspiration. The second two were socio-economic in nature: the estimation of agriculture yield and production, and an agent-based model simulating household dynamics in dryland agroecosystems.  The first and third analyses were relatively simple, in that the variable(s) of interest were mapped onto landcover using empirical relationships. The second and fourth relied on more complex numerical methods, where landcover was one of several variables needed to run each model. For the simpler analyses, we examined how results were influenced by map aggregation, while for the more complex cases, our assessments were confined to each numerical model's standard output resolution.

\vspace{-0.5 cm}
\section{Map quality}
\subsection{Bias and accuracy}

Our reference map indicated that crop fields covered nearly 10\% of the total study area in the 2009-2011 time period. Subtracting each test map from the reference maps created pixel-wise residuals, where negative and positive values respectively represent overestimates and underestimates by the test map (Fig. 1A).

%\end{wrapfigure}

%\begin{figure*}[htp]
%\centering
%    \subfigure[random caption 1]{\includegraphics[scale=0.49]{figures/bias_map.pdf}}\quad
%     \subfigure[random caption 1]{\includegraphics[scale=0.49]{figures/bias_map.pdf}}
%\end{figure*}
%\linenumbers
The most pronounced errors were in the MODIS and GlobCover maps, which showed large positive residuals in the center of the country where cropland is most concentrated (blue areas in Fig. 1A), and negative residuals (red areas) along the eastern and northern margins.
These patterns translated into substantial map bias (Fig. 1B), with GlobCover and MODIS mean error exceeding 45\% and 25\% respectively at 1 km resolution, meaning that each map tends to underestimate cropland by that amount at that resolution. This bias declined with each level of map aggregation, being reduced to nearly 15\% for GlobCover and 5\% for MODIS at 100 km. The magnitude of mean absolute error (MAE) was somewhat higher in all cases. The GeoWiki map, in contrast, was the least biased overall, showing just a ~7\% bias at 1 km and near 0 for all other scales of aggregation, although its accuracy (23\% MAE) was only half as good as SA-LC's at 1 km (11\% MAE), which despite its uniform overestimation bias (Fig. 1A) was the most accurate map at aggregation scales $<10$km. Above this, GeoWiki became slightly more accurate, having $<$5\% MAE at 100 km resolution. The reason GeoWiki had relatively poor accuracy at 1 km resolution was due to the high heterogeneous error pattern, which traded between positive and negative residuals over short distances, thereby inflating MAE at this scale.

\subsection{Error and landcover density}
Given that landcover characteristics can influence landcover map quality \cite{see_improved_2015, estes_platform_2015, debats_generalized_2015}, we set out to elucidate the relationship between cropland configuration and map error.  We first calculated the average 1 km reference cropland percentage within each of South Africa's 354 magisterial districts (the finest administrative unit, averaging 3,445 km$^2$; SI), which yielded a landscape-scaled metric of characteristic cropland density, and similarly calculated the district-wise MAE for each test map.

>>>>

%% PNAStwoS.tex
%% Sample file to use for PNAS articles prepared in LaTeX
%% For two column PNAS articles
%% Version1: Apr 15, 2008
%% Version2: Oct 04, 2013
%
%% BASIC CLASS FILE
%\documentclass{pnas}
%
%% ADDITIONAL OPTIONAL STYLE FILES Font specification
%
%\usepackage{pnastwoF}
%\usepackage{cite}
%\usepackage[numbers,round,sort,compress]{natbib}
%\bibpunct{(}{)}{,}{n}{,}{,}  % https://xianblog.wordpress.com/tag/natbib/ (allows natbib with PNAS)
%\usepackage{lineno}
%\setlength\linenumbersep{2pt}
%\usepackage{placeins}
%\usepackage{float}
%\restylefloat*{figure}
%\usepackage{dblfloatfix}  % http://tex.stackexchange.com/questions/235623/placing-a-figure-in-the-bottom-of-a-page-spanning-the-two-columns-of-an-ieee-doc 
%\usepackage{subfigure}
%\usepackage{wrapfig}
%\usepackage{caption}
%\setlength\belowcaptionskip{-2cm}
%\setlength{\textfloatsep}{5pt plus 1.0pt minus 2.0pt}
%\usepackage{multirow}
%\setlength{\parskip}{1cm plus4mm minus3mm}
%\usepackage{makecell}
%\usepackage{boldline}
%\setcellgapes{3pt}
%\usepackage[flushleft]{threeparttable}
%
%% OPTIONAL MACRO DEFINITIONS
%\def\s{\sigma}
%%%%%%%%%%%%
%% For PNAS Only:
%\url{www.pnas.org/cgi/doi/10.1073/pnas.0709640104}
%\copyrightyear{2008}
%\issuedate{Issue Date}
%\volume{Volume}
%\issuenumber{Issue Number}
%\setcounter{page}{2687} %Set page number here if desired
%%%%%%%%%%%%
%
%
%
\documentclass[a4paper]{article}
\usepackage[affil-it]{authblk}

\usepackage{graphicx}
\usepackage{tabularx}
\usepackage{epstopdf}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{amsthm}
%\usepackage{endfloat}
\usepackage{float}
\usepackage[sort,compress]{natbib}
%\bibpunct{[}{]}{,}{n}{,}{,}  % https://xianblog.wordpress.com/tag/natbib/ (allows natbib with PNAS)
\usepackage{endnotes}
\usepackage{setspace}
\usepackage{verbatim}
\usepackage[left=2.5cm, right=2.5cm, bottom=2cm, top=2cm]{geometry}
\usepackage{times}
\usepackage{helvet}
\usepackage{courier}
%\usepackage{mathtime}
\usepackage{bm}
\usepackage{url}
%\usepackage{babel}
\usepackage{dcolumn}
\usepackage{multirow}
\usepackage{makecell}
\usepackage{boldline}
\setcellgapes{3pt}
\usepackage[flushleft]{threeparttable}
\usepackage{wrapfig}
\newcommand{\citetapos}[1]{\citeauthor{#1}'s \citeyearpar{#1}}
%%%%%

\usepackage[
  breaklinks=true,
  colorlinks=true,
  linkcolor=blue,anchorcolor=blue,
  citecolor=blue,filecolor=blue,
  menucolor=blue,pagecolor=blue,
  urlcolor=blue]{hyperref}

% Let's add todonotes and comments:
%\usepackage[draft]{todonotes}
%
%% Select what to do with command \comment:  
%% \newcommand{\comment}[1]{}  %comment not showed
%\newcommand{\comment}[1]
%{{[\bfseries \color{blue} #1]}} %comment showed

\usepackage{lineno}
\usepackage{float}
\usepackage[anythingbreaks]{breakurl}

%\begin{document}
%\hspace*{\fill}

@@

%\author{Lyndon Estes{\affil{1}{Woodrow Wilson School, Princeton University, Princeton, NJ USA}\affil{2}{Civil and Environmental Engineering, Princeton University, Princeton, NJ USA}},
%Peng Chen\affil{3}{Indiana University, Bloomington, IN USA},
%Stephanie Debats\affil{2}{},
%Tom Evans\affil{3}{},
%Stefanus Ferreira\affil{4}{GeoTerraImage, Pretoria, RSA}{},
%Gabrielle Ragazzo\affil{5}{Ecology and Evolutionary Biology, Princeton University, Princeton, NJ USA},
%Justin Sheffield\affil{2}{},
%Adam Wolf\affil{5}{},
%Eric Wood\affil{2}{},
%\and
%Kelly Caylor\affil{2}{}}

\author[1,2]{Lyndon Estes}
\author[3]{Peng Chen}
\author[1]{Stephanie Debats}
\author[3]{Tom Evans}
\author[4]{Stefanus Ferreira}
\author[5]{Gabrielle Ragazzo}
\author[1]{Justin Sheffield}
\author[5]{Adam Wolf}
\author[1]{Eric Wood}
\author[1]{Kelly Caylor}

\affil[2]{Woodrow Wilson School, Princeton University, Princeton, NJ USA}
\affil[1]{Civil and Environmental Engineering, Princeton University, Princeton, NJ USA}
\affil[3]{Indiana University, Bloomington, IN USA}
\affil[4]{GeoTerraImage, Pretoria, RSA}
\affil[5]{Ecology and Evolutionary Biology, Princeton University, Princeton, NJ USA}

@@

\begin{document}
\maketitle

%\begin{article}
\begin{abstract}
{Research into global environmental change is often built upon landcover maps, which can have substantial errors. How these errors impact our understanding of global change processes and associated policies is thus of great interest, but is difficult to fully quantify because we lack spatially comprehensive reference data, particularly in the World's most rapidly developing regions. We used a high quality, national-scale cropland map to assess bias and accuracy in current generation landcover maps and four examples of ``downstream'' (landcover-dependent) analyses, two related to physical processes (estimates of vegetative carbon stocks and evapotranspiration) and two to socio-economic processes (gridded crop production estimates and agent-based simulation of household food security), at resolutions ranging from 1-100 km. We found that the landcover maps have substantial errors below 25 km resolution, with substantial biases and inaccuracies, particularly from maps derived from coarse resolution sensors. In some cases these error metrics remain high even when maps are aggregated to 100 km. These errors can be substantially amplified in downstream analyses (e.g. the carbon and crop production analyses), but in cases where the values of different cover types that underpin the analysis are similar (e.g. evapotranspiration estimates) the results can be relatively insensitive. To avoid these errors, and thus minimize the risk of misunderstanding global change processes or misinforming policy, substantial map aggregation of landcover maps from their base resolution is often needed. We provide several guidelines that can help landcover map users select appropriate datasets and  aggregation scales under different use cases.

@@

%\keywords{landcover | bias | remote sensing | agriculture | crop yield | harvested area | carbon | agent-based model | landscape}

@@

\section*{Introduction}
The functioning of the Earth System is fundamentally connected to the characteristics of landcover \citep{lambin_modelling_1997}, the physical constituents of the terrestrial surface. Human endeavors are strongly governed by and shape landcover, and the characteristics of landcover govern many climate and biogeochemical processes \citep{lambin_dynamics_2003}. Our increasing modification of the Earth's surface \citep{lambin_dynamics_2003} means that socioeconomic and physical processes increasingly interact through landcover. To fully understand these processes and the nature of global change, an accurate understanding of the nature and distribution of landcover is essential.

This importance is understood by a growing number of social, economic, and natural scientists, who are using landcover data to advance understanding of food security \citep{lark_cropland_2015,wright_recent_2013, licker_mind_2010}, carbon cycling \citep{asner_high-resolution_2010, gaveau_major_2014}, biodiversity loss \citep{newbold_global_2015, luoto_predicting_2004}, demographic shifts \citep{linard_assessing_2010}, and other important facets of global change.

The value of the insights resulting from such studies depends upon the veracity of their underlying landcover data, much as a house requires a solid foundation in order to remain standing. Unfortunately, the evidence to date indicates that much of our understanding of global change is built on shaky foundations. The reason for this is that landcover data can only practically be derived from satellite imaging, which has several important constraints, related to both physics and analytics, that propagate mapping errors. First, in many regions the cover types of interest are smaller than the sensor resolution \citep[e.g. smallholder's farms][]{jain_mapping_2013,debats_generalized_2016,ozdogan_resolution_2006}, or spectrally indistinct from neighboring covers \citep{sweeney_mapping_2015}, which are factors that increase mapping complexity \citep{yu_meta-discoveries_2014}. Second, the act of defining a cover class can cause error, in that selected classes may have highly diverse spectral properties (e.g. agriculture \citep{debats_generalized_2016,estes_platform_2015}) and thus be difficult for the classifier to distinguish, while discretizing a continuous cover type (e.g. dividing a forest into different canopy cover classes) can also promote classification error, particularly near class boundaries \citep{foody_status_2002}, as well as confusion about the actual extent of the cover type \citep{sexton_conservation_2015}. Furthermore, class definitions often vary between maps, complicating inter-comparison \citep{kuemmerle_challenges_2013}. Third, landcover map series are often used to detect changes \citep[e.g.][]{gross_monitoring_2013}, or may use within-class seasonality as a means for discriminating between classes (e.g. phenology of cropping versus surrounding vegetation \citep{sweeney_mapping_2015}), but seasonal variability and landcover changes can be easily confused. Given these multiple sources of error, landcover maps are often inaccurate at finer scales and disagree widely between products, particularly in the world's most rapidly developing regions \citep{estes_projected_2013, fritz_comparison_2010, fritz_cropland_2011}. These errors limit our ability to obtain granular, mechanistic understanding of global change processes.

These problems with landcover products are known \citep{fritz_comparison_2010, fritz_cropland_2011, see_improved_2015, fritz_mapping_2015,verburg_challenges_2011}, and there are a variety of map improvement efforts underway \citep[e.g.][]{fritz_geo-wiki:_2012, estes_platform_2015}. What remains an open question is exactly how much the maps researchers typically use deviate from actual landcover, and how this in turn impacts our understanding of global change processes. Fully answering this question depends on having comprehensive, spatially representative ground truth data, which are typically unavailable for Africa and other rapidly changing regions \citep{see_improved_2015}, which have also been fairly sparsely mapped \citep{yu_meta-discoveries_2014}. Our current understanding of map accuracy over such areas is therefore based on a handful of maps \citep{yu_meta-discoveries_2014}, which were assessed using either bottom-up tests made with a relatively small number of ground truth points, or from top-down "sanity checks" made in comparison to aggregated survey data.  In terms of the former, there are well-developed, best-practice guidelines for selecting and using ground-truth samples to robustly quantify map error, but these are not commonly applied by the broader community of landcover map users (often due to lack of data), which makes it challenging to compare how error varies within and between maps \citep{stehman_global_2012, olofsson_making_2013,olofsson_good_2014,foody_status_2002}. The latter form of assessment is made at coarser administrative levels relative to reported statistics \citep[e.g.][]{fritz_comparison_2010}, which can also be suspect \citep{carletto_emperor_2013,fao_action_2013}.

Since it is difficult to fully quantify landcover map errors, it is even more challenging to gauge their impact on downstream analyses, where there is substantial risk of error amplification\citep{kuemmerle_challenges_2013}. There has been some work examining how map error influences climate simulations \citep{ge_impacts_2007}, agricultural land use patterns \citep{schmit_limitations_2006}, carbon flux measurements \citep{quaife_impact_2008}, human population estimates \citep{linard_assessing_2010}, and species distribution models \citep{tuanmu_global_2014}, but these either use simulated landcover errors \citep{ge_impacts_2007} or compare relevant differences in estimates between different satellite-derived landcover maps \citep{linard_assessing_2010, quaife_impact_2008,tuanmu_global_2014}. One exception is a Belgian study \citep{schmit_limitations_2006} that used ground-collected farm parcel data to assess how landcover errors bias measurements of agricultural land use patterns, but the extent of this study was fairly small and the validation data were discontiguous.

Fortunately, the recent, explosive growth in public and private initiatives to develop new Earth observing capabilities, which range from small drones\footnote{e.g. 3DRobotics, DJIA} to new high resolution satellite arrays\footnote{Planet Labs, Skybox} and better mapping methods \citep{fritz_geo-wiki:_2012,tuanmu_global_2014,estes_projected_2013,debats_generalized_2016}, are finally providing the means to more comprehensively interrogate the accuracy and biases in the landcover products that have become commonplace in global change research--and which are often used to make policy decisions \citep{searchinger_high_2015}.

In this study, we take advantage of this recent growth in high resolution data to address the call to more thoroughly assess errors and in landcover maps \citep{kuemmerle_challenges_2013, olofsson_good_2014,olofsson_global_2012}, and further examine how these errors might impact our understanding of change in the world's most dynamic regions.  We use a unique, high accuracy landcover map of South African crop fields to conduct a spatially comprehensive, bottom-up quantification of error in several widely used landcover maps, and how these errors can propagate through downstream studies investigating both the physical and socioeconomic components of global change. Our objective is to provide scientists and policy-makers who use landcover data an up-to-date assessment of their appropriate uses and limitations.

\vspace{-0.5 cm}
\section*{Materials and Methods}
\subsection*{Overview of study area and reference data}
In the late 2000s, the South African government commissioned a cropland map that was made by manually interpreting and digitizing fields visible within high resolution satellite imagery \citep{fourie_better_2009}. The resulting vectorized field boundaries provide highly accurate data on field sizes and distribution for the period 2009-2011. Measured in terms of its ability to distinguish crop fields from other cover types, this dataset is 97\% percent accurate at 1 km resolution, with user's accuracies of 94\% and 98\% for the cropland and non-cropland classes, and producer's accuracies of 84 and 99\% \citep[][SI]{estes_platform_2016}.

This dataset is particularly valuable because South Africa represents nearly 6\% of sub-Saharan Africa's (SSA) area, which is a region that is poised to undergo rapid agricultural expansion \citep{searchinger_high_2015}, yet is notably lacking trustworthy maps of existing agriculture land \citep{fritz_comparison_2010}. Moreover, South Africa's large agricultural sector represents the diversity of farming systems found throughout SSA, ranging from large commercial operations to smallholder farms \citep{hardy_rainfed_2011,estes_using_2014}.

\subsubsection*{Assessment of landcover map accuracy}
We used the vectorize field boundaries as a reference layer for evaluating four landcover products representative of the type commonly used in global change studies. The first was South Africa's own 30 m resolution 2009 National Landcover map \citep[SA-LC][]{sanbi_national_2009}, which is typical of the higher-resolution, Landsat-based maps that are typically available only for individual countries \citep[e.g.][]{fry_completion_2009}. The second and third were respectively the 300 m GlobCover 2009 \citep{arino_global_2012} and 500 m resolution 2011 MODIS Landcover products, which are widely used global-scale products \citep[e.g.][]{gross_monitoring_2013, shackelford_conservation_2015}. The fourth dataset was the new 1 km GeoWiki hybrid-fusion cropland map for Africa \citep{fritz_mapping_2015}, which incorporates the first three datasets and represents the current state-of-the-art in landcover mapping.  Since GeoWiki provides a continuous bounded value of landcover (percent cropland cover), which cannot be feasibly converted to a categorical value, we converted all other datasets, including the reference vectors, into comparable 1 km gridded percent cropland estimates.

To develop the cropland reference map (see SI for accuracy assessment details), we intersected the original field boundary vector maps with a 1 km grid, and calculated the percent of each cell occupied by fields to create the gridded cropland reference map. In the resulting grid, we masked out areas classified as communal farmland, because only their outer perimeters were digitized \citep{fourie_better_2009}, which posed a risk of overestimating cropland extent. We also excluded the permanent tree crop class, to avoid confusion with landcover maps that exclude these in their cropland classes. For the same reason, we used other high resolution landcover maps for selected provinces of South Africa to mask out commercially afforested areas and sugarcane plantations, which were also not included in the reference map. The resulting masked reference grid covered 90\% of South Africa (1,081,000 km$^2$), of which 104,304 km$^2$ was cropland.

We extracted the cropland classes from SA-LC, MODIS, and GlobCover, and converted these into percent cropland estimates at 1 km resolution. Both MODIS and GlobCover had mixed pixel classes, thus, following the approach used in creating the GeoWiki maps \citep{fritz_mapping_2015} we used the upper, mean, and lower cropland estimates from these classes to produce three versions of the gridded percentages. We used the mean map for the main analysis, but estimated error variability using all three versions (SI).

We first evaluated the quality of the landcover product-derived percent cropland maps (hereafter referred to as the ``test maps''). Instead of the standard confusion matrix-based accuracy assessment metrics \citep{olofsson_good_2014,olofsson_making_2013}, which apply to categorical landcover maps, we assessed the bias and accuracy of the test maps based on the gridded residuals that resulted when each test map was subtracted from the reference map. Here bias is the mean residual value, weighted by the density of reference cropland, and accuracy is the mean of the absolute values of residuals (also weighted by cropland density), thus accuracy is inversely related to this value. We calculated these metrics for the original 1 km resolution, and for maps that were further aggregated to 5, 10, 25, 50, and 100 km, in order to evaluate how bias and accuracy changes with scale. For these aggregated maps, we applied a further weight in calculating error metrics, the number of pixels contributing to each aggregated pixel, to prevent pixels close to national boundaries or where non-target cover types were masked out from having outsize influence on the statistics.

We also assessed how landcover pattern impacts map performance by modeling the correlation between map accuracy and cropland density.  To evaluate this relationship, we used magisterial district boundaries (n=354, mean area=3,445 km$^2$; SI) to provide a landscape-scaled unit for calculating characteristic cover density. We filtered out pixels with $<$0.05\% (0.5 ha) cropland, to prevent the much larger areas of non-agricultural districts from dominating the signal, extracted the absolute values of test map errors and corresponding reference cropland percentages, and calculated their district-wide means. The relationship between mean absolute error (response) and cropland density (predictor) was then modeled using a generalized additive model \citep{hastie_generalized_1990}, with each district weighted by its number of agricultural pixels \citep{wood_mgcv:_2001}. To account for potential spatial autocorrelation, we fit a two-dimensional smoothing spline to the coordinates of each district's centroid.

\subsection*{Impact of map error on downstream analyses}
We then used these maps to conduct four further analyses typical of global change research. The first two related to physical processes, namely the calculation of vegetative carbon stocks and the simulation of evapotranspiration. The second two were socio-economic in nature: the estimation of agriculture yield and production, and an agent-based model simulating household dynamics in dryland agroecosystems.  The first and third analyses were relatively simple, in that the variable(s) of interest were mapped onto landcover using empirical relationships. The second and fourth relied on more complex numerical methods, where landcover was one of several variables needed to run each model. For the simpler analyses, we examined how results were influenced by map aggregation, while for the more complex cases, our assessments were confined to each numerical model's standard output resolution.

\subsubsection*{Biophysical analyses}
\emph{Estimating vegetative carbon stocks: }
To understand the carbon cycle and climate forcing due to land use change, it is important to have accurate, high resolution maps of vegetative carbon stocks \citep[][]{searchinger_high_2015}. One widely used vegetative carbon dataset is that of \citep{ruesch_new_2008}, who mapped estimated carbon density values for different vegetation types to the classes of a global landcover product. The resulting data were intended to provide a baseline for climate policy by the Intergovernmental Panel on Climate Change (IPCC), as well as input to other land use and biogeochemical analyses \citep{ruesch_new_2008}.

%\vspace{-1 cm}
We followed this method to create vegetative carbon maps for South Africa. Since our cropland percentage map provided no information on surrounding cover, we developed several variants representing potential surrounding cover types by assigning the average carbon densities of five biomes (forest, secondary forest, shrubland, grassland, and sparse vegetation \citep{ruesch_new_2008}) to the non-cropland fraction of our maps. These hypothetical maps represented the range in potential carbon densities, and allowed us to investigate how carbon estimates can vary as a function of i) test map errors and ii) the properties of neighboring cover types. We multiplied cropland densities by cropland fractions and added these to each of the five other densities multiplied by the residual non-cropland fractions to create five different carbon density maps. We aggregated each carbon map up to the five coarser resolutions for scaling comparisons (SI).

To assess carbon estimation error, we subtracted test map-derived carbon maps from those based on the reference map, and calculated bias and accuracy scores using the method described for the cropland maps (see previous section).
